Numerical simulation of the shallow water circular hydraulic jump

### 2020-09-14 numerics_paper/numerical_chj.tex
Update numerical_chj.tex
Change the phrase "The value of $\theta$ is calculated as the residual" by the correct one "The value of $\theta$ is calculated as the 2-norm of the residual"
Add PyClaw simulation results.

\usepackage{amsmath}
\usepackage{graphicx}

@@

\author{
    David I. Ketcheson \and
    Manuel Quezada de Luna \and
    David Batista
}
\begin{document}

@@

where $Fr_{u}$ is the Froude number perpendicular to the cell face and $\varepsilon$, $\alpha$, and $\beta$ are parameters that have to be chosen appropriately. For the shallow water equations, authors in \cite{kemm2014note} propose to use $\varepsilon=10^{-3}$ and $\alpha=\beta=\frac{1}{3}$. The value of $\theta$ is calculated as the residual in the Rankine-Hugoniot condition for the shear wave.\\

@@

>>>>

\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{geometry}
\usepackage{float}
\usepackage{subfig}

@@

\author{
    David I. Ketcheson \and
    Manuel Quezada de Luna
}
\begin{document}

@@

where $Fr_{u}$ is the Froude number perpendicular to the cell face and $\varepsilon$, $\alpha$, and $\beta$ are parameters that have to be chosen appropriately. For the shallow water equations, authors in \cite{kemm2014note} propose to use $\varepsilon=10^{-3}$ and $\alpha=\beta=\frac{1}{3}$. The value of $\theta$ is calculated as the 2-norm of the residual in the Rankine-Hugoniot condition for the shear wave.\\

@@

To demonstrate the range of qualitative solutions that can arise when using
different numerical methods, we apply several schemes to the following setup.
We take
\begin{align*}
    r_0 & = 0.1 & h_0 & = 0.5 & u_0 & = 0.75 \\
    g & = 1 & h_\infty & = 0.15 & F_\infty & = 0.1
\end{align*}
where $r_0, h_0, u_0$ are the jet radius and the depth and velocity at the outer
edge of the jet, respectively; $h_\infty$ and $F_\infty$ are the depth
and Froude number (infinitely) far away from the jet.  The domain is $[-1,1]^2$
and boundary conditions are imposed based on the 1D steady state solution with
the chosen jet and far-field conditions.  The bathymetry is flat and friction is
neglected.  The initial data is chosen to be far from the equilibrium and
consists of a uniform depth of $h=0.15$ and zero velocity.
Results using a $1000\times 1000$ grid are shown in Figure \ref{fig:hlle_vs_roe}.  
We see that by $t=4$, the HLLE solution has already reached the equilibrium
and remains there indefinitely.  In contrast, the Roe solution is highly
unsteady with a very non-circular jump containing cusps that tend to form
near the four points where the jump is aligned with a grid axis.

Based on numerous tests, we have found that results for other initial data
are very similar to those shown here, although of course the details
of the Roe solution differ because the solution is chaotic.

\begin{figure}
  \begin{centering}
    \subfloat[HLLE Riemann solver, t=4]{
      \includegraphics[width=0.45\textwidth]{figures/hlle_1000_t4.eps} 
    }
    \subfloat[Roe Riemann solver with entropy fix, t=4]{
      \includegraphics[width=0.45\textwidth]{figures/roe_1000_t4.eps}
    }

    \subfloat[HLLE Riemann solver, t=20\label{fig:hlle}]{
      \includegraphics[width=0.45\textwidth]{figures/hlle_1000_t20.eps} 
    }
    \subfloat[Roe Riemann solver with entropy fix, t=20\label{fig:roe}]{
      \includegraphics[width=0.45\textwidth]{figures/roe_1000_t20.eps}
    }
  \end{centering}
  \caption{
    Comparison of PyClaw solutions using the classic Clawpack discretization
    with HLLE and Roe Riemann solvers.\label{fig:hlle_vs_roe}}
\end{figure}

Next we apply Kemm's HLLEMCC solver to the same scenario; results are shown
in Figure \ref{fig:hllemcc}.  Results at early times are very similar to
those for the Roe solver.  The cusps that appear have a stronger tendency
to dissipate and then reform, as indicated in the figure at $t=20$.
Nevertheless, they continue to appear indefinitely, based on longer-time
simulations.

\begin{figure}
  \begin{centering}
    \subfloat[HLLEMCC Riemann solver, t=4]{
      \includegraphics[width=0.45\textwidth]{figures/hllemcc_1000_t4.eps} 
    }
    \subfloat[HLLEMCC Riemann solver, t=20]{
      \includegraphics[width=0.45\textwidth]{figures/hllemcc_1000_t20.eps}
    }
  \end{centering}
  \caption{
    PyClaw solutions using the classic Clawpack discretization
    with HLLEMCC Riemann solver.\label{fig:hllemcc}}
\end{figure}
